---
title: "A Decade of Sunspots: 2009 -2019"
subtitle: "(Using Kanzelhöhe Observatory ISN)"
author: "David J Jackson"
date: "1/30/2020"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(prophet)
library(scales)
#
# Total days 27,639
kanzel <- fread("./../db/kh_spots.csv")
kanzel$Ymd <- as.Date(kanzel$Ymd)
kanzel$Year <- year(kanzel$Ymd)
kanzel$Month <- month(kanzel$Ymd)
kanzel$S <- kanzel$s_n + kanzel$s_s
kanzel$G <- kanzel$g_n + kanzel$g_s
kanzel$Cts <- ifelse(kanzel$R ==0,0,1)
kanzel$Yes <- ifelse(kanzel$R >=1,"Yes","No")
kanzel<-kanzel[,.(Ymd,Year,Month,g_n,s_n,g_s,s_s,R,R_n,R_s,S,G,Cts,Yes)]
df <- kanzel %>% filter(Ymd >="2009-01-01" & Ymd <="2019-12-31")
```

# International Sunspot Number(ISN/Rn)
"The daily (mean) ISN data comes from ,Kanzelhöhe Observatory(http://cesar.kso.ac.at/spots/kh_spotsquery.php) University  of Graz, Aust."

"The Wolf number (also known as the International sunspot number, relative sunspot number, or Zürich number) is a quantity that measures the number of sunspots and groups of sunspots present on the surface of the Sun.", https://en.wikipedia.org/wiki/Wolf_number

```{r pressure, echo=FALSE}
ggplot(data=df,aes(x=Ymd,y=R)) +geom_line() +geom_smooth() + 
ggtitle("Total ISN(Rn) Value: 2009 - 2019")
```

# Kanzel Total by Year: Groups and Spots

```{r summaries, echo=FALSE}
df.group <- df %>% select(Year,Groups) %>% group_by(Year) %>% summarize(Group = sum(Groups))
df.spots <- df %>% select(Year,Spots) %>% group_by(Year) %>% summarize(Spots = sum(Spots))
```
```{r plot1,echo=FALSE}
ggplot(data=df.spots,aes(x=Year,y=Spots)) + geom_col() + ggtitle("Kanzel Total Sunspots by Year")
ggplot(data=df.group,aes(x=Year,y=Group)) + geom_col() + ggtitle("Kanzel Total Sunspot Groups by Year")
```
```{r echo=FALSE}
ggplot(data=df,aes(Groups)) +geom_histogram(binwidth=2) + 
  facet_wrap(~Year,nrow =4,ncol=4,scale="free_y") + 
ggtitle("Histogram of Sunspot Groups by Year: 2009 - 2019") +xlab("Year by Year") + ylab("Total Groups")
```
```{r echo=FALSE}
ggplot(data=df,aes(Spots)) +geom_histogram(binwidth=20) + facet_wrap(~Year,nrow =4,ncol=4,scale="free_y") + 
ggtitle("Histogram of Total Sunspots by Year: 2009 - 2019") +xlab("Year by Year") + ylab("Total Spots")
```

# Summary of Days with amd with Out Spots by Year
```{r stackeding days, echo=FALSE}
df %>% group_by(Year,Yes) %>%  summarise(Days= n()) %>% spread(Year,Days)
df %>% group_by(Year,Yes) %>%  summarise(Days= n()) %>%
ggplot(aes(x=Year,y=Days,fill=Yes)) + geom_bar(stat="identity")
```

\newpage
# Cycle 25 and Beyond: Predicting Monthly ISN

```{r predict1, echo=FALSE}
df.pred <- kanzel %>% select(Ymd,Spots)
colnames(df.pred) <- c("ds","y")
m <- prophet(seasonality.mode="multiplicative")
m <- add_seasonality(m, name="cycle_11year", period=365.25 * 11,fourier.order=5)
m <- fit.prophet(m, df.pred)
future <- make_future_dataframe(m,periods=8000,freq="day")
forecast <- predict(m, future)
plot(m,forecast) +xlab("Year of Prediction") + ylab("Total Sunspots")
```
```{r cycle25, echo=FALSE}
fcast <- forecast %>% select(ds,yhat,yhat_lower,yhat_upper) %>% filter(ds>="2020-01-01" & ds <="2025-12-31")
ggplot(data=fcast,aes(x=ds,y=yhat,col="Predicted")) +geom_line() +xlab("Year of Prediction") + ylab("Total Sunspots") +geom_smooth(method="loess",aes(col="loess")) +
ggtitle("Solar Cycle 25 and Beyond: 2020 - 2026")

```
```{r}
library(mice)
df_mice <- kanzel %>% select(Ymd,S,G,R)
```
```{r Import SIDC Data}
sidc <- fread("../db/SN_d_tot_V2.0.csv")
colnames(sidc) <- c("Year","Month","Day", "Fdate","R", "Sd","Obs" ,"Defin"  )
sidc$Ymd <- as.Date(paste(sidc$Year, sidc$Month, sidc$Day, sep = "-"))
sidc$S <- NA
sidc$G <- NA
sidc <- sidc %>% filter(Ymd >="1944-05-29") %>% select(Ymd,S,G,R)

```


